Non-kissing complexes and tau-tilting for gentle algebras

### 2018-07-13 firstPaper/corrections/replyToReferee.txt

We thank the referee for their helpful comments.  We have tried to implement them
all.  Below are our replies to each comment.  We have also made a few additional
changes, which we first list.

(a) In Notation 1.50, the equalities were wrong and have been corrected.

(b) We have added some details (in a new Proposition 2.24) explaining why the 
    non-kissing facets are finite. In doing so, we had to slightly reorganise 
    Subsection 2.2.2 (new Lemma 2.23 was previously part of the proof of 
    Proposition (renumbered) 2.28).

@@

* Figure 17: $Q_{pa}$ as above why is an example given a notation-like name?

  Reply:

* Figure 18: crossing walks have not been introduced

  Reply: We have replaced "crossing" by "non-kissing and having a common substring".  
         The term "crossing walks" appeared nowhere else in the paper.

@@

* remark 3.72: do you mean ``Figure 36 (bottom)''?

  Reply:

A note about notation. This article uses a quite creative notation set. For
example the blossoming quiver has a little flower as exponent and the operation
of blinding is represented by two half circle to suggest blinkers. I am not sure
I am happy about these choices but I do not have better suggestions to offer.

  Reply: While we agree that the notation set is non-conventional, we believe
         that it fits well with our chosen terminology.  We thus respectfully 
         propose to keep the notation as it is.

### 2018-07-13 firstPaper/firstPaperLong/nonKissingComplex.tex
Some more referee's changes to first paper.

\captionsetup{width=1.4\textwidth}
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{table}
{
	\hspace{1cm}
	\begin{tabular}{|c|c|c|}
    \hline
    String modules on~$\bar Q$ &
    Walks on~$\bar Q\blossom$  &
    Reference
    \\ \hline \hline
    almost positive strings (Def.~\ref{def:stringsBands}) &
    bending walks (Def.~\ref{def:straightBended}) &
    Fig.~\ref{exm:exmBijectionStringsWalks1} \& \ref{fig:exmAuslanderReitenQuiver}
    \\ \hline
    ?\footnotemark[1] &
    straight walks (Def.~\ref{def:straightBended}) &
    Fig.~\ref{exm:exmBijectionStringsWalks1} \& \ref{fig:exmAuslanderReitenQuiver}
    \\ \hline
    \multiLinesBox{$P(v)$ and $P(v)[1]$ \\ $A=kQ/I$ and~$A[1]$ \\ ~} &
    \multiLinesBox{$v_\peak$ and $v_\deep$ \\ $F_\peak$ and~$F_\deep$ \\ (Exm.~\ref{example: Fdeep Fpeak} and Def.~\ref{def: deep walk})} &
    Lem.~\ref{lem:bijections-STTC-NKC}
    \\ \hline
    ?\footnotemark[2] &
    countercurrent order~$\prec_\alpha$ &
    ---
    \\ \hline
    $\tau$-compatibility (Def.~\ref{def: tau-compatibility}) &
    non-kissing (Def.~\ref{def: kissing}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    support~$\tau$-tilting object (Def.~\ref{def: tau-rigid and stautilt}) &
    maximal non-kissing collection of walks &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    support~$\tau$-tilting complex~$\tTC$ (Def.~\ref{def: stautilt complex}) &
    reduced non-kissing complex~$\RNKC$ (Def.~\ref{def: nKc}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    \multiLinesBox{(left) mutation of support \\ $\tau$-tilting objects (Thm.~\ref{thm: mutation of stautilts}~\cite{AdachiIyamaReiten})} &
    \multiLinesBox{(increasing) flip \\ (Prop.~\ref{prop:flip} and Def.~\ref{def:flip} \& \ref{def:increasingFlip})} &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    lattice of torsion classes &
    non-kissing lattice~$\NKL$ (Thm.~\ref{thm:lattice}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    bricks &
    \multiLinesBox{distinguishable strings (Def.~\ref{def:distinguishedSubstring})} &
    Prop.~\ref{prop:characterizationDistinguishableStrings}
    \\ \hline
    collections of pairwise Hom-orthogonal bricks &
    non-friendly complex (Def.~\ref{def: friendly}) &
    Prop.~\ref{prop:characterizationDistinguishableStrings}
    \\ \hline
    \multiLinesBox{bijection between indecomposable~$\tau$-rigid representations and bricks (Thm. 6.1 in~\cite{DemonetIyamaJasso})} &
    \multiLinesBox{bijection between non-kissing walks that are not peak walks and distinguishable strings (Prop.~\ref{prop:bijectionDistinguishableStringsWalks})} &
    \multiLinesBox{Conjectural \\ cf.~Prop.~\ref{prop:bijectionDistinguishableStringsJI}}[2cm]
    \\ \hline
    \multiLinesBox{Bongartz cocompletion of a~$\tau$-rigid indecomposable representation $N=M(\rho)$ whose associated brick is~$M(\sigma)$. Moreover $\sigma = \distinguishedWalk{\omega(\rho)}{F}$} &
    \multiLinesBox{$F=\eta\left(\closure{\Sigma_\bottom(\sigma)}\right)$, where $\sigma$ is a distinguishable string} &
    Rem.~\ref{rem: Bongartz}
    \\ \hline
    Bongartz completion &
    $\eta\left(\closure{\Sigma_\top(\sigma)}\right)$ &
    Rem.~\ref{rem: Bongartz}
    \\ \hline
    $\b{g}$-vector or index of a module (Def.~\ref{definition: g-vector of representation}) &
    $\b{g}$-vector of a walk (Def.~\ref{def: g-vectors for walks}) &
    Rem.~\ref{rem: g-vectors coincide}
    \\ \hline
    $\dim \Hom{A} \big( M(\sigma), \tau M(\sigma') \big)$ (Sec.~\ref{subsec:tautilting}) &
    kissing number~$\kappa \big( \omega(\sigma), \omega(\sigma') \big)$ (Def.~\ref{def: kissing number}) &
    Rem.~\ref{rem:KNvsTau}
    \\ \hline
    \end{tabular}
	\hspace{1cm}
}
{Dictionary between algebraic notions on the support $\tau$-tilting complex and combinatorial notions on the non-kissing complex.}
{table:dictionary}
\captionsetup{width=\textwidth}

@@

\footnotetext[1]{The straight walks should morally correspond to some projective-injective objects in an exact category. We plan to further investigate the situation depicted in Figure~\ref{fig:exmAuslanderReitenQuiver}.}
\footnotetext[2]{We do not know of an algebraic interpretation of the countercurrent order yet. We nonetheless included it in the table above since we think it might be of interest to find one.}

@@

\begin{figure}[b]
	\capstart
	\centerline{\includegraphics[scale=1]{kissingCrossing}}
	\caption{A schematic representation of two kissing walks (left) and two crossing walks~(right): $\omega$~kisses~$\omega'$ (left) while~$\mu$~crosses~$\nu$ (right). For the crossing walks, we have~$\mu \prec_\gamma \nu$ while~$\nu \prec_\delta \mu$.}
	\label{fig:kissingCrossing}
\end{figure}

@@

\begin{proof}
\enlargethispage{.3cm}
Throughout the proof, we denote~$S \eqdef \zeta(F') \ssm \{\sigma\}$ and~$S' \eqdef \zeta(F')$.
Consider an increasing flip~$F \to F'$ with~$F \ssm \{\omega\} = F' \ssm \{\omega'\}$ and supported by~$\sigma = \distinguishedString{w}{F} = \distinguishedString{w'}{F'}$.
We prove that~$S$ is biclosed:
\begin{description}
\item[closed] Otherwise, there exist bottom substrings~$\sigma_1, \dots, \sigma_\ell$ of walks~$\omega_1, \dots, \omega_\ell$ of~$F'$, arrows~$\alpha_1, \dots, \alpha_{\ell-1} \in Q_1$ and signs~$\varepsilon_1, \dots, \varepsilon_{\ell-1} \in \{-1,1\}$ such that~$\sigma = \sigma_1 \alpha_1^{\varepsilon_1} \dots \alpha_{\ell-1}^{\varepsilon_{\ell-1}} \sigma_\ell$.
Assume that there is~$k \in [\ell-1]$ such that~${\varepsilon_k = 1}$ and choose~$k$ minimal with that property.
Then~$\omega_k \ne \omega'$ and~$\omega$ has two outgoing arrows while~$\omega_k$ has two incoming arrows at the endpoints of~$\sigma_k$, so that~$\sigma_k \in \Sigma_\top(\omega) \cap \Sigma_\bottom(\omega_k)$, contradicting the non-kissingness of~$F$.
We obtain a similar contradiction assuming that there is~$k \in [\ell-1]$ such that~$\varepsilon_k = -1$ and choosing~$k$ maximal with that property.
This forces~$\ell = 1$, so that~$\sigma \in S$, a contradiction.

@@

\begin{corollary}
\label{coro:finalLatticeQuotient}
If~$\bar Q$ is a gentle bound quiver whose non-kissing complex~$\RNKC$ is finite, then the non-kissing oriented flip graph~$\NKG$ is the Hasse diagram of the lattice quotient~$\Bicl{\bar Q}/{\equiv}$ of the congruence-uniform lattice of biclosed sets of strings by the lattice congruence~$\equiv$.
\end{corollary}

@@

Conversely, for a non-self-kissing walk~$\omega \in \NKWalks^\pm(\bar Q)$ not in the peak facet, we denote by~$\sigma_\bottom(\omega)$ the inclusion minimal string in~$\set{\sigma \in \Sigma_\bottom(\omega)}{\closure{\Sigma_\bottom(\omega)} = \closure{\Sigma_\bottom(\sigma)}}$.
\end{definition}

@@

In order to describe the canonical join-representations in the next section, we now completely describe the join-irreducible elements in~$\NKL$.
We first consider the following join-irreducible elements of~$\Bicl{\bar Q}$.

\begin{lemma}
\label{lem:exmBiclosedJI}
For any distinguishable string~$\sigma \in \strings^\pm(\bar Q)$, the biclosed set~$\closure{\Sigma_\bottom(\sigma)}$ is join-irreducible in~$\Bicl{\bar Q}$.
\end{lemma}

@@

\begin{theorem}
The following assertions are equivalent for a subset~$\Sigma$ of~$\distinguishableStrings^\pm(\bar Q)$:
\begin{enumerate}[(i)]
\item $\Sigma$ is pairwise non-friendly.
\item $\ji(\Sigma) \eqdef \set{\ji(\sigma)}{\sigma \in \Sigma}$ is the canonical join-representation of an element in~$\NKL$.
\item $\Sigma$ is the descent set of a non-kissing facet~$F \in \NKC$.
\end{enumerate}
Thus, the canonical join complex of~$\NKL$ is isomorphic to the non-friendly complex~$\NFC$.
\end{theorem}

@@

\begin{proposition}
\label{prop:gvectorscvectorsDualBases}
For a non-kissing facet~$F \in \RNKC$, the set of $\b{g}$-vectors~$\gvectors{F}$ and the set of $\b{c}$-vectors~$\cvectors{F}$ form dual bases. 
\end{proposition}

>>>>

\captionsetup{width=1.4\textwidth}
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{table}
{
	\hspace{1cm}
	\begin{tabular}{|c|c|c|}
    \hline
    String modules on~$\bar Q$ &
    Walks on~$\bar Q\blossom$  &
    Reference
    \\ \hline \hline
    almost positive strings (Def.~\ref{def:stringsBands}) &
    bending walks (Def.~\ref{def:straightBended}) &
    Fig.~\ref{exm:exmBijectionStringsWalks1} \& \ref{fig:exmAuslanderReitenQuiver}
    \\ \hline
    ?\footnotemark[1] &
    straight walks (Def.~\ref{def:straightBended}) &
    Fig.~\ref{exm:exmBijectionStringsWalks1} \& \ref{fig:exmAuslanderReitenQuiver}
    \\ \hline
    \multiLinesBox{$P(v)$ and $P(v)[1]$ \\ $A=kQ/I$ and~$A[1]$ \\ ~} &
    \multiLinesBox{$v_\peak$ and $v_\deep$ \\ $F_\peak$ and~$F_\deep$ \\ (Exm.~\ref{example: Fdeep Fpeak} and Def.~\ref{def: deep walk})} &
    Lem.~\ref{lem:bijections-STTC-NKC}
    \\ \hline
    ?\footnotemark[2] &
    countercurrent order~$\prec_\alpha$ &
    ---
    \\ \hline
    $\tau$-compatibility (Def.~\ref{def: tau-compatibility}) &
    non-kissing (Def.~\ref{def: kissing}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    support~$\tau$-tilting object (Def.~\ref{def: tau-rigid and stautilt}) &
    maximal non-kissing collection of walks &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    support~$\tau$-tilting complex~$\tTC$ (Def.~\ref{def: stautilt complex}) &
    reduced non-kissing complex~$\RNKC$ (Def.~\ref{def: nKc}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    \multiLinesBox{(left) mutation of support \\ $\tau$-tilting objects (Thm.~\ref{thm: mutation of stautilts}~\cite{AdachiIyamaReiten})} &
    \multiLinesBox{(increasing) flip \\ (Prop.~\ref{prop:flip} and Def.~\ref{def:flip} \& \ref{def:increasingFlip})} &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    lattice of torsion classes &
    non-kissing lattice~$\NKL$ (Thm.~\ref{thm:lattice}) &
    Thm.~\ref{thm:nkc/sttiltc}
    \\ \hline
    bricks &
    \multiLinesBox{distinguishable strings (Def.~\ref{def:distinguishedSubstring})} &
    Prop.~\ref{prop:characterizationDistinguishableStrings}
    \\ \hline
    collections of pairwise Hom-orthogonal bricks &
    non-friendly complex (Def.~\ref{def: friendly}) &
    Prop.~\ref{prop:characterizationDistinguishableStrings}
    \\ \hline
    \multiLinesBox{bijection between indecomposable~$\tau$-rigid representations and bricks (Thm. 6.1 in~\cite{DemonetIyamaJasso})} &
    \multiLinesBox{bijection between non-kissing walks that are not peak walks and distinguishable strings (Prop.~\ref{prop:bijectionDistinguishableStringsWalks})} &
    \multiLinesBox{Conjectural \\ cf.~Prop.~\ref{prop:bijectionDistinguishableStringsJI}}[2cm]
    \\ \hline
    \multiLinesBox{Bongartz cocompletion of a~$\tau$-rigid indecomposable representation $N=M(\rho)$ whose associated brick is~$M(\sigma)$. Moreover $\sigma = \distinguishedWalk{\omega(\rho)}{F}$} &
    \multiLinesBox{$F=\eta\left(\closure{\Sigma_\bottom(\sigma)}\right)$, where $\sigma$ is a distinguishable string} &
    Rem.~\ref{rem: Bongartz}
    \\ \hline
    Bongartz completion &
    $\eta\left(\closure{\Sigma_\top(\sigma)}\right)$ &
    Rem.~\ref{rem: Bongartz}
    \\ \hline
    $\b{g}$-vector or index of a module (Def.~\ref{definition: g-vector of representation}) &
    $\b{g}$-vector of a walk (Def.~\ref{def: g-vectors for walks}) &
    Rem.~\ref{rem: g-vectors coincide}
    \\ \hline
    $\dim \Hom{A} \big( M(\sigma), \tau M(\sigma') \big)$ (Sec.~\ref{subsec:tautilting}) &
    kissing number~$\kappa \big( \omega(\sigma), \omega(\sigma') \big)$ (Def.~\ref{def: kissing number}) &
    Rem.~\ref{rem:KNvsTau}
    \\ \hline
    \end{tabular}
	\hspace{1cm}	
}
{Dictionary between algebraic notions on the support $\tau$-tilting complex and combinatorial notions on the non-kissing complex.
 \footnotetext[1]{The straight walks should morally correspond to some projective-injective objects in an exact category. We plan to further investigate the situation depicted in Figure~\ref{fig:exmAuslanderReitenQuiver}.}
 \footnotetext[2]{We do not know of an algebraic interpretation of the countercurrent order yet. We nonetheless included it in the table above since we think it might be of interest to find one.}
}
{table:dictionary}

\captionsetup{width=\textwidth}

@@

% \footnotetext[1]{The straight walks should morally correspond to some projective-injective objects in an exact category. We plan to further investigate the situation depicted in Figure~\ref{fig:exmAuslanderReitenQuiver}.}
% \footnotetext[2]{We do not know of an algebraic interpretation of the countercurrent order yet. We nonetheless included it in the table above since we think it might be of interest to find one.}

@@

\begin{figure}[b]
	\capstart
	\centerline{\includegraphics[scale=1]{kissingCrossing}}
	\caption{A schematic representation of two kissing walks (left) and two non-kissing walks~(right) having a common substring: $\omega$~kisses~$\omega'$ (left) while~$\mu$~does not kiss~$\nu$ (right). On the right, we have~$\mu \prec_\gamma \nu$ and~$\nu \prec_\delta \mu$.}
	\label{fig:kissingCrossing}
\end{figure}

@@

\begin{proof}
\enlargethispage{.3cm}
Throughout the proof, we denote~$S \eqdef \zeta(F') \ssm \{\sigma\}$ and~$S' \eqdef \zeta(F')$.
Consider an increasing flip~$F \to F'$ with~$F \ssm \{\omega\} = F' \ssm \{\omega'\}$ and supported by~$\sigma = \distinguishedString{w}{F} = \distinguishedString{w'}{F'}$.
We prove that~$S$ is biclosed:
\begin{description}
\item[closed] Otherwise, there exist bottom substrings~$\sigma_1, \dots, \sigma_\ell$ of respective walks~$\omega_1, \dots, \omega_\ell$ of~$F'$, arrows~$\alpha_1, \dots, \alpha_{\ell-1} \in Q_1$ and signs~$\varepsilon_1, \dots, \varepsilon_{\ell-1} \in \{-1,1\}$ such that~$\sigma = \sigma_1 \alpha_1^{\varepsilon_1} \dots \alpha_{\ell-1}^{\varepsilon_{\ell-1}} \sigma_\ell$.
Assume that there is~$k \in [\ell-1]$ such that~${\varepsilon_k = 1}$ and choose~$k$ minimal with that property.
Then~$\omega_k \ne \omega'$ and~$\omega$ has two outgoing arrows while~$\omega_k$ has two incoming arrows at the endpoints of~$\sigma_k$, so that~$\sigma_k \in \Sigma_\top(\omega) \cap \Sigma_\bottom(\omega_k)$, contradicting the non-kissingness of~$F$.
We obtain a similar contradiction assuming that there is~$k \in [\ell-1]$ such that~$\varepsilon_k = -1$ and choosing~$k$ maximal with that property.
This forces~$\ell = 1$, so that~$\sigma \in S$, a contradiction.

@@

\begin{corollary}
\label{coro:finalLatticeQuotient}
If~$\bar Q$ is a gentle bound quiver such that the non-kissing complex~$\RNKC$ is finite, then the non-kissing oriented flip graph~$\NKG$ is the Hasse diagram of the lattice quotient~$\Bicl{\bar Q}/{\equiv}$ of the congruence-uniform lattice of biclosed sets of strings by the lattice congruence~$\equiv$.
\end{corollary}

@@

Conversely, for a non-self-kissing walk~$\omega \in \NKWalks^\pm(\bar Q)$ not in the peak facet, the inclusion minimal string in the set~$\set{\sigma \in \Sigma_\bottom(\omega)}{\closure{\Sigma_\bottom(\omega)} = \closure{\Sigma_\bottom(\sigma)}}$ is denoted by~$\sigma_\bottom(\omega)$.
\end{definition}

@@

In order to describe the canonical join-representations in the next section, we now completely describe the join-irreducible elements in~$\NKL$.
Let us first consider the following join-irreducible elements of~$\Bicl{\bar Q}$.

\begin{lemma}
\label{lem:exmBiclosedJI}
Let~$\sigma \in \strings^\pm(\bar Q)$ be a distinguishable string. Then the biclosed set~$\closure{\Sigma_\bottom(\sigma)}$ is join-irreducible in~$\Bicl{\bar Q}$.
\end{lemma}

@@

\begin{theorem}
The following assertions are equivalent for a subset~$\Sigma$ of~$\distinguishableStrings^\pm(\bar Q)$:
\begin{enumerate}[(i)]
\item $\Sigma$ is pairwise non-friendly.
\item the set~$\ji(\Sigma) \eqdef \set{\ji(\sigma)}{\sigma \in \Sigma}$ is the canonical join-representation of an element in~$\NKL$.
\item $\Sigma$ is the descent set of a non-kissing facet~$F \in \NKC$.
\end{enumerate}
Thus, the canonical join complex of~$\NKL$ is isomorphic to the non-friendly complex~$\NFC$.
\end{theorem}

@@

\begin{proposition}
\label{prop:gvectorscvectorsDualBases}
For a non-kissing facet~$F \in \RNKC$, the set~$\gvectors{F}$ of~$\b{g}$-vectors and the set~$\cvectors{F}$ of~$\b{c}$-vectors form dual bases. 
\end{proposition}

### 2018-07-13 firstPaper/firstPaperShort/nonKissingComplex.tex
Some more referee's changes to first paper.

\begin{theorem*}
For any gentle bound quiver~$\bar Q$ whose non-kissing complex is finite, the non-kissing lattice is congruence-uniform.
In fact, the non-kissing lattice is both a lattice quotient and a sublattice of a lattice of biclosed sets on strings in~$\bar Q$.
\end{theorem*}

\noindent
It is then natural to study the canonical join complex of the non-kissing lattice, which we call the non-friendly complex following~\cite{GarverMcConville-grid}.
This complex corresponds to the classical non-crossing partitions in Catalan combinatorics.

>>>>

\begin{theorem*}
For any gentle bound quiver~$\bar Q$ whose non-kissing complex is finite, the non-kissing lattice is congruence-uniform.
In fact, the non-kissing lattice is both a lattice quotient and a sublattice of a lattice of biclosed sets on strings in~$\bar Q$.
\end{theorem*}
\footnotetext[1]{The straight walks should morally correspond to some projective-injective objects in an exact category. We plan to further investigate the situation depicted in Figure~\ref{fig:exmAuslanderReitenQuiver}.}
\footnotetext[2]{We do not know of an algebraic interpretation of the countercurrent order yet. We nonetheless included it in the table above since we think it might be of interest to find one.}
\noindent
It is then natural to study the canonical join complex of the non-kissing lattice, which we call the non-friendly complex following~\cite{GarverMcConville-grid}.
This complex corresponds to the classical non-crossing partitions in Catalan combinatorics.
